GPs spent on average 25\% of their working day handling clinical paperwork and updating Electronic Patient Records (EPR) systems. This application aims to ease the stress and time taken by medical professionals by harnessing an Ambient Voice Technology (AVT) approach to record, transcribe, diarise, and summarise medical consultations to reduce time spent during and after consultations taking notes and follow-up reports. Furthermore, the appraisals feature allows doctors to generate appraisal entries by supplementing the summarisation with additional appraisal notes and tags. These entries can be viewed in the history and searched for by name or tag. Finally, by enabling the doctor to upload medical guidance documentation, the application will search for relevant guidelines to enable the GP to explore and cite medical advice based on the conversation. By engineering this project as an offline application, we expand the use case to in-home visits and ensure that all sensitive data is safe due to not requiring an internet connection for any part of the application. This application aims to reduce the time taken recording notes and inputting data into EPR systems by providing concrete transcripts, summarisations and medical advice based on the consultation.

An iterative approach allowed for flexibility in the project's development, allowing for the feedback from supervisors and healthcare professionals to be quickly integrated into the current solution. This project built on a summer project using an Intel Gaudi server running a LLM to summarise patient-doctor consultations. This was extended further to move the application onto the Windows platform, running entirely offline with the additional transcription pipeline and a RAG model for the search for relevant medical guidelines.

This project harnesses the power of Intel's AIPCs by using OpenVINO to quantise machine learning models to reduce the inference time resulting in an increase in speed and decrease in application size, producing fast and accurate summarisations of medical conversations, with medical advice provided if required.