\chapter{Testing}
% Describe your testing strategy (unit, functional, acceptance testing and
% how they are carried out). How were test cases selected?
% – Examples of specific tests and how they were carried out (e.g., using
% mock objects to break dependencies). Focus on the interesting cases.
% – A summary of the test results and what coverage was achieved.
% Detailed test reports should appear in the appendix, if they add useful
% information or you want to demonstrate the kinds of tests and
% coverage achieved.
% If your project requires substantial evaluation of data and results,
% evaluation of algorithms, or other forms of testing that are not code-based,
% then adapt this chapter to suit.
% This chapter will typically be 2-4 pages in length but could be more
% depending on the depth of testing done. If you need to do a detailed
% evaluation for a more mathematical or theory-based project, then this
% chapter could well be more substantial.

\section{Strategy}
The testing strategy for this project uses requirement driven testing to systematically evaluate the system against the technical, functional and non-functional requirements. This was conducted through unit, UI, performance, accuracy testing and user acceptance testing in iterative stages through the development process. A test execution matrix was created based on the requirements to ensure all requirements map to a test case found in Appendix \ref{app:test_matrix}

\section{Functional Testing}
Functional testing evaluates the application's ability to execute its core workflows and manage data correctly. This is conducted with unit tests for individual components and pipelines using Visual Studio's integrated test runner. The \texttt{gtest} library was chosen due to the integration within the Visual Studio environment allowing for quick configuration and running of the tests. When creating the unit tests for the first prototype of the application (transcription/summarisation pipeline), the architecture for the program prevented mocking external libraries. All external library and model calls were extracted to an interface to allow for dependency injection. The tests then implement a mock interface to inject into the function for testing. The resulting design required minimal code changes and loss of functionality or performance change while allowing for proper mocking of back-end functionalities. An example is shown below.

\begin{verbatim}
// SummarisationEngine.h
struct ISummarisationBackend {
    virtual ~ISummarisationBackend() = default;
    virtual void LoadModel(const std::filesystem::path& modelPath) = 0;
    virtual std::string Generate(const std::string& prompt, int maxNewTokens
                                                            , float temperature) = 0;
};
// SummarisationEngine.cpp
void SummarisationEngine::SetBackendForTesting(std::shared_ptr<ISummarisationBackend> backend) {
    m_testBackend = std::move(backend);
}
void SummarisationEngine::loadModel() {
    std::filesystem::path modelPath = Helpers::GetModelPath("Med42-int4");
    if (m_testBackend) {
        m_testBackend->LoadModel(modelPath);
        return;
    }

    m_model = new ov::genai::LLMPipeline(modelPath, "GPU");
}

\end{verbatim}

This strategy was used throughout the program, allowing for the creation of mock classes for robust unit testing. The full list of unit tests found in Appendix \ref{app:unittest} with the results in Appendix \ref{app:unittest_passing}

\section{UI Testing}
To test the user interface, all buttons and fields were navigated to ensure the flow and data persistence for correctness. A full table of the tests and results can be found in appendix \ref{app:ui_test}

\section{Performance Testing}
Performance testing was conducted to verify that the application meets the latency and responsiveness constraints defined in requirements NF1-4. All tests were executed on the target Intel AI PC.

\begin{table}[H]
\centering
\begin{tabular}{|l|p{6cm}|l|l|}
\hline
\textbf{Req ID} & \textbf{Performance Metric} & \textbf{Target Constraint} & \textbf{Observed Result} \\ \hline
NF1 & Application startup and complete ML model loading into memory & $< 60$ seconds & Med42: $20$s, other models: $4$s \\ \hline
NF2 & Summary generation completion after a 15-minute consultation & $< 90$ seconds & $73$ seconds \\ \hline
NF3 & Semantic search and guidance retrieval over uploaded documents & $< 3$ seconds & $\approx 2$ seconds \\ \hline
NF4 & UI Responsiveness during heavy background processing & No freezing/hangs & None observed \\ \hline
\end{tabular}
\caption{Summary of performance testing results against non-functional requirements.}
\label{tab:performance_results}
\end{table}

Informal testing involved monitoring the task manager to record the GPU and CPU useage and memory consumption. During transcription, the CPU usage maintained 40\% with 22GB of RAM usage. Once the Med42-v2 model was loaded, VRAM usage maintained 6.5GB with 100\% usage at inference time.


\subsubsection{Summarisation Inference (NF2)}
To test the NF2 requirement, a 15 minute mock conversation was conducted with a friend. By recording the time taken for the conversation and inference, we found the result of around 73 seconds for a 15 minute conversation, within the required 90 seconds.

\subsubsection{Retrieval Latency (NF3)}
This requirement was tested by uploading 3 medical PDFs and pressing "view guidance". On an average of 10 runs, the time taken between pressing "view guidance" and the UI displaying the highlighted guidance within the PDF averaged around 2 seconds.

\subsubsection{UI Responsiveness (NF4)}
Throughout all UI testing, the user interface remained responsive with no noticeable delay

\subsection{Accuracy Testing}
\subsubsection{Summarisation}
After selecting Med42-v2 as the summarisation model, the LLM testing framework in section \ref{di:test_med42} was re-ran with the \texttt{rescale\_with\_baseline\=False} to provide an individual benchmark over a comparative benchmark. The results provided in \ref{fig:med42_benchmark} demonstrate this is a strong candidate precision score of 88\%.

\begin{figure}[H]
    \centering
    \begin{tikzpicture}
        \begin{axis}[
            ybar,
            title={Performance Metrics for Med42-int4},
            ylabel={Score},
            symbolic x coords={F1, Precision, Recall},
            xtick=data,
            ymin=0.0, ymax=1.0, % Zoomed in slightly to show the differences better
            nodes near coords, % Displays the numbers on top of the bars
            nodes near coords align={vertical},
            bar width=25pt,
            enlarge x limits=0.3, % Adds space on the left and right sides
            width=8cm,
            height=6cm
        ]
        
        \addplot[fill=blue!40] coordinates {
            (F1, 0.888) 
            (Precision, 0.882) 
            (Recall, 0.895)
        };
        
        \end{axis}
    \end{tikzpicture}
    \label{fig:med42_benchmark}
    \caption{Evaluation metrics for the Med42-int4 model.}
\end{figure}

\subsubsection
General performance of the Whisper-Medium model has been validated by numerous studies finding a word error rate of 5.86\% to 9.4\% for English speakers in a quiet controlled environment \cite{pulikodan2025approach,chang2026llm}. Therefore the accuracy of the transcription in this clinical setting was tested on medical terminology. A glossary of medical terminology published by the European Medicines Agency \cite{ema_medical_terms_simplifier} was used as the ground truth set. This was conducted by running the application and speaking 4 randomly selected words for each starting letter of the alphabet. Once completed, the transcription was viewed and compared against the glossary. To analyse these results, each word was placed into a category of correct, minor inaccuracy (punctuation or American English) and major inaccuracy.

\begin{table}[htbp]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Count} & \textbf{Correct} & \textbf{Inaccuracy} & \textbf{Incorrect} \\
\hline
104 & 93 & 8 & 3 \\
\hline
\end{tabular}
\caption{Transcription results with Whisper-Medium on 104 medical terminology samples}
\label{tab:results}
\end{table}

Generally, the transcription model has a large enough vocabulary to accurately transcribe medical terminology with a 7\% word error rate. Common inaccuracies involve incorrect hyphens, American English spelling, and incorrect capitalisation. The incorrectly transcribed words included syncope, fissure and tophi miss-transcribed to sync up, Fisher and trophy. This is likely due to the transcribed words having higher weights within the model due to their popularity compared to the niche medical terminology. The full results can be examined in Appendix \ref{app:testing_transcription_results}.


\subsubsection{Guidelines Retrieval}
Due to the custom retrieval pipeline implementation, testing with a ground truth PDF dataset is unfeasible. Without a found ground truth set, no quantitative testing or results can be generated. Qualitative testing showed the retrieved guidance to be correct, however the usability was highly dependent on the specificity of the guidance PDFs uploaded. General documents laying out a disease progression or advice for trivial diseases provided little valuable insight. 

\section{UI Testing}
All buttons and data entry fields were tested throughout the application, ensuring state transitions and all behaviour was as expected. A full list of test cases and results is in Appendix \ref{app:UI_test}.

\section{User Acceptance Testing}
Throughout the project, 5 demonstrations and 2 trials of the application were conducted with Dr Atia Rafiq and Dr Tazeen Ahmed, providing a formative evaluation of the application. At a UCL hosted student showcase, I trialled the application (without the guidelines retrieval) by mocking a consultation with Dr. Tazeen Ahmed where she used the application as if she was conducting a real consultation. She was able to use the application with little guidance and the consultation was successfully carried out resulting in a successful user acceptance test(TODO requirement). She expressed her enthusiasm with the project and commented that this application could save her lots of time taking notes and reviewing consultations throughout the day. She further commented that the appraisals system could be highly beneficial for training GPs. She also provided feedback regarding the prompt, noting the informal tone, suggesting a "the patient says" and "the doctor suggested" format which was integrated after the event. Unfortunately due to a cyber attack on the St. Georges Hospital, a full mock consultation and planned formal user acceptance testing could not be completed.

On a video call with Dr Atia Rafiq, I demonstrated the application (without the guidelines retrieval and appraisals). She noted that the user interface looked easy to use and had an appealing design and that the application could be very useful to her and her trainee GPs. She added a further recommendation to create the appraisals component to allow note taking and commentary on conversations to demonstrate learning that was conducted post-consultation.




